---
title: "TRANSPORTATION PROBLEM"
author: "SANIA FATIMA"
date: "2023-10-10"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
***
SUMMARY:
Objective Function:Minimize TC=622x11+614x12+630x13+
                               641x21+645x22+649x23
                               
Subject to Constraints[Supply]:x11+x12+x13>=100
                               x21+x22+x23>=120
                      [Demand]:x11+x21>=80
                               x12+x22>=60
                               x13+x23>=70
        
Now, Subject to Non-Negativity Constraints: xij=0,where i=1,2 and j=1,2,3                       
 ***                       

#Loading Packages
```{r}
library(Matrix,warn.conflicts = FALSE)
library(lpSolve,warn.conflicts = FALSE)
```
  
#Building the matrix of the given problem
```{r}
transmatrix <- matrix(c(22,14,30,600,100,
                        16,20,24,625,120,
                        80,60,70,"-","210/220"),ncol=5,nrow=3,byrow = TRUE)
colnames(transmatrix) <- c("Warehouse1","Warehouse2","Warehouse3",
                           "Production Cost","Production Capacity")
rownames(transmatrix) <- c("PlantA","PlantB","Monthly Demand")
transmatrix <- as.table(transmatrix)
transmatrix
```
***
```{r}
new.transmatrix <- matrix(c(622,614,630,0,100,
                            641,645,649,0,120,
                            80,60,70,10,220),ncol=5,nrow=3,byrow=TRUE)

colnames(new.transmatrix) <- c("Warehouse1","Warehouse2","Warehouse3",
                               "Dummy","Production Capacity")
rownames(new.transmatrix) <- c("PlantA","PlantB","Monthly Demand")
transmatrix <- as.table(new.transmatrix)
transmatrix

```

***
#The balanced values of the problem will be satisfied by this relation. The cost Matrix which we have created is shown below:

```{r}
costs<-matrix(c(622,614,630,0,
                641,645,649,0),nrow = 2, byrow = TRUE)
costs
```
#The values of the matrix of the row's Production Capacity side are as follows: 
```{r}
row.rhs<-c(100,120)
row.signs<-rep("<=",2)
```

#Here, we used the double variable 10 at the end to determines the dummy variable and also we used the values of the matrix from column's side Production Capacity are as follows:
```{r}
col.rhs<-c(80,60,70,10)
col.signs<-rep(">=",4)
```

#In this chunk, we are going to use the LP Transport Command to run the code
```{r}
lptrans<-lp.transport(costs,"min",row.signs,row.rhs,col.signs,col.rhs )
lptrans$solution
lptrans$objval
```
***
Conclusion:From the Above Code,we can conclude that our result is Z=132790.The result for each variable of both plants are as follows:
    60x12 which is warehouse2 from plantA
    40x13 which is warehouse3 from plantA
    80x21 which is warehouse1 from plantB
    30x23 which is warehouse3 from plantB
    10x24 which is the 4rth variable is called as the "throw away variable".
    